\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-09-24 papers/methods/paper.tex
Updating with feedback from Dan

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{Steve}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}
\newcommand{\green}{\color{forestgreen}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

@@

% Abstract of the paper
\begin{abstract}
We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects. This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improving treatment of Malmquist bias, whilst increasing numerical efficiency. Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery. We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample. SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and being more simulation independent than previous analyses.
\end{abstract}
% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the width of this distribution is called the intrinsic dispersion of the supernova population and the hat denotes an observed variable.

@@

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size. Whilst binning data does result in some loss of information, recent works tested against simulations show that this information is not cosmologically significant \citep{Scolnic2016, Kessler2017}.

@@

Whilst there are a large number of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.

@@

Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. As the SALT2 model does not include intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$. These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

@@

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by $\sigma_{M_B}$ without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.

As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature). The $\kappa$ parameters are tightly constrained with priors as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$. We subject them to strong Cauchy priors centered on zero and with width $0.002$, where the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulting in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion.  Each survey has its own set of $\kappa$ parameters. As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.

Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient. We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.

@@

\subsubsection{Type Ia population}

The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude. We follow the prior work of {\rubin} and model the colour population as an independent redshift-dependent skew normal for each survey. For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift. Each set of nodes are anchored to equally spaced to span the minimum to maximum redshift of each survey. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment. We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.

@@

For future use with photometrically classified surveys, we include a simplistic outlier population model. We follows {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works. However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. The probability of each supernovae falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section. For the spectroscopic survey, we set this to unity. For the photometric proof of concepts we provide an accurate probability estimate. Further investigation on the effect of inaccurate estimates we leave for future improvements during the analysis of the DES photometric sample.

@@

We tested two models with varying $\beta$ against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$. See Section \ref{sec:simdes} for details on simulation generation. We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects. This degeneracy resulted in a significant bias in recovered cosmology. As $\alpha$ and $\beta$ are constants in the SALT2 model, assuming varying $\alpha$ or $\beta$ is only self consistent if the same variations are used in the light curve fitting process. As this variation is not present in the DES spectroscopic sample's light curve fitting process, and because of the recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses. As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.

@@

Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.

@@

In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based. We note that there was no significant difference in the selection effect of Malmquist bias between using the {\gten} or {\celeven} scatter model compared to the fit uncertainty.

@@

Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}. The parameter bounds are listed in Table \ref{tab:bulk_summary}. Our recovered results are all within $0.5\sigma$ of the true values, however some bias does emerge.

We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics. This is confirmed by replacing the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. We then also test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$. From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.

Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data. It is difficult to do whilst accounting for correlations with population and scatter model. This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently. Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.

>>>>

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{Steve}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\definecolor{forestgreen}{rgb}{0.13, 0.75, 0.13}
\newcommand{\green}{\color{forestgreen}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

@@

% Abstract of the paper
\begin{abstract}
We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects. This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improving treatment of Malmquist bias, whilst increasing numerical efficiency. Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery. We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample. SNANA simulations can recover the input cosmology {\green between $0$ and $0.5\sigma$ bias depending on the intrinsic scatter model} for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and increasing simulation independent over previous analyses. {\green Future development will focus on increasing model agnosticism towards intrinsic scatter model.} {\red Have left out explicit references for `previous analyses' as I dont want to bloat the abstract. Its talked about in the paper, so is this alright?}
\end{abstract}
% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. {\green This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}.} {\red  Ive just swapped out references here as this isnt a critical point for my analysis to drill down on.} The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable. {\red Removed confusing statement on intrinsic dispersion.}

@@

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size. Whilst binning data does result in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

@@

Whilst there are a large number of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$. {\red Dan, you mention underselling the prize here. Any suggestions on an appropriate sell, so to speak?}

@@

Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. As the SALT2 model does not include {\green full treatment of} intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$. These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

@@

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the {\green width of the underlying magnitude population (discussed in Section \ref{sec:underlying})} without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion. {\green Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.}

As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature). The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$. {\green We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$. We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.} As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.

Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, {\green by modelling the dispersion as a multivariate Gaussian}, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient. We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations. This extra dispersion is modelled with redshift independence. {\red I'm not too sure what you mean about telling a story here. I've tried to change the wording to put it back in line with the method, is that sufficient? Also, I did not try redshift dependent extra scatter on $m_B$.}

@@

\subsubsection{Type Ia population} 
\label{sec:underlying}
The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude. We follow the prior work of {\rubin} and model the colour population as an independent redshift-dependent skew normal for each survey. For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift. Each set of nodes are anchored to equally spaced to span the minimum to maximum redshift of each survey. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment. We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.

@@

For future use with photometrically classified surveys, we include a simplistic outlier population model. We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works. However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. The probability of each supernovae falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section. For the spectroscopic survey, we set this to unity. For the photometric proof of concepts we provide an accurate probability estimate. Further investigation on the effect of inaccurate estimates we leave for future improvements during the analysis of the DES photometric sample.

@@

We tested two models with varying $\beta$ against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$. See Section \ref{sec:simdes} for details on simulation generation. We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects. This degeneracy resulted in a significant bias in recovered cosmology. {\red Removed comments about varying alpha and beta.} {\green Due to the }recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses. As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.

@@

Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population {\green \citep{Conley2011, Betoule2014}}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.

@@

In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based. {\green We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the selection effect of Malmquist bias between the two models.}

@@

Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}. The parameter bounds are listed in Table \ref{tab:bulk_summary}. {\green Our recovered results are all within $0.5\sigma$ of the true values of $\Omega_m = 0.3$, $w=-1$.}

We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics. {\green To confirm this, we run two tests. The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.} From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.

Unfortunately, adding extra fit parameters to allow for shifting observables washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data. It is difficult to do whilst accounting for correlations with population and scatter model. This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently. Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
